One-stage point transect distance sampling using iterated integrated nested Laplace approximations

### 2019-12-04 paper-v2.tex
add communication of results section

Since CV values are influenced both by sampling effort and the magnitude of the estimated intensity means other methods we also map the lower and upper quantiles corresponding to a 95\% credible interval for each prediction cell (\autoref{fig:intensity-quantiles}).
\begin{figure}[!t]
	\includegraphics[scale=0.5]{figures/intensity_quantiles.png}
	\caption{The predicted posterior intensity summarised in three plots by the mean (upper),  0.025 quantile (lower left) and 0.975 quantile (lower right) for each prediction grid cell.  All three plots use a common colour scale}
	\label{fig:intensity-quantiles}
\end{figure}

@@

% POSTERIOR MATERN CORRELATION HERE
Add posterior matern correlation function

\begin{figure}[h]
	\includegraphics[scale=0.6]{figures/matern_posterior.png}
	\caption{Posterior Mat\'ern correlation function}
	\label{fig:half-normal}
\end{figure}

@@

@@

\clearpage %figure out figure placements later
\section{Model Evaluation}

\begin{itemize}
	\item mini lit review - communicating uncertainty in spatial models w random effects
\end{itemize}

Discuss wisely

>>>>

We also map the lower and upper quantiles corresponding to a 95\% credible interval for each prediction cell (\autoref{fig:intensity-quantiles}).
\begin{figure}[h]
	\includegraphics[scale=0.5]{figures/intensity_quantiles.png}
	\caption{The predicted posterior intensity summarised in three plots by the mean (upper),  0.025 quantile (lower left) and 0.975 quantile (lower right) for each prediction grid cell.  All three plots use a common colour scale}
	\label{fig:intensity-quantiles}
\end{figure}

@@

The posterior Mat\'ern correlation function is shown in \autoref{fig:matern-posterior}, the grey regions show the 95\% credible interval. 
\begin{figure}[h]
	\includegraphics[scale=0.6]{figures/matern_posterior.png}
	\caption{Posterior Mat\'ern correlation function}
	\label{fig:matern-posterior}
\end{figure}
Note that this correlation function shows positive correlation even at distances of 10,000 metres, wider than the full length of the study region (approximately 8,000 metres).  This is the correlation on the log-intensity scale.  [[Can I make a better "range of correlation" plot for the posterior intensity process?]]

@@

% Add table of estimated model parameters
% Add 95% CI figures for N
%% EVAL SPDE AND LGCP POSTERIOR SAMPLING OF barplot k function

\clearpage %sort out figure placements later

\section{Communication of Results}

% This is not quite right - model communication?  
% in Results add a section evaluating SPDE using boxplot k function

\begin{itemize}
	\item mini lit review - communicating uncertainty in spatial models w random effects
	\item IDEA:  lots of papers don't even both themselves with uncertainty.  Could cite a bunch of recent ones, noting the search term in web of knowledge.
\end{itemize}

The results presented above are consistent with a more traditional two-stage model on binned count data.  Our presentation of the results are broadly consistent with approaches taken in the species distribution modelling literature, although coefficient of variation maps (\autoref{fig:intensity-mean-cv}) and quantile maps (\autoref{fig:intensity-quantiles}) are not always provided.  We note a worrying tendency in the literature to only provide maps of point estimates without any accompanying communication of uncertainty.  Where uncertainty maps are provided they tend to be maps of posterior CV, standard deviation (see for example [[CITE THEM]]) and/or posterior or bootstrapped quantiles (see for example [[CITE]]).  Our intention in this section is to highlight limitations with our above communication of the results and discuss possible remedies.

Whilst summary maps of the posterior intensity process are useful, they can mask certain properties of the fitted model. The mapped posterior mean, for example, will be qualitatively smoother than realisations from the posterior.\autoref{fig:intensity-realizations} shows three such realisations.  Note that each of them has finer-grained spatial structure than is shown in the posterior mean in \autoref{fig:intensity-mean-cv}.  One interpretation of this is that animals will tend to be more clustered than might be expected if we only looked at the posterior mean.

\begin{figure}
	\includegraphics[scale=0.35]{figures/intensity_realized.png}
	\caption{Three realizations of the posterior intensity field}
	\label{fig:intensity-realizations}
\end{figure}

During model evaluation we strongly recommend producing plots of realisations as these will more qualitatively resemble spatial structure of the observed data.

More concerning are the problems with the CV, standard deviation and quantile maps.  The CV map is useful but problematic to interpret in regions of low estimated intensity.  If the posterior standard deviation is relatively consistent across the survey region, CV values will be higher in regions of low predicted intensity.  The observed patterns in such maps will just highlight regions of low intensity, not necessarily higher uncertainty.  This is further confounded in the case with spatially varying posterior standard deviation making the CV map hard to interpet.

The quantile maps also suffer from a potential difficulty in interpretation. The temptation is to interpet the maps as showing possible intensity surfaces that could have produced the observed data.  Presenting these quantiles side by side in a single map obscures the fact that it is vanishingly unlikely for all prediction values to simultaneously achieve the 0.025 or 0.975 quantile.  We believe most non-statistically trained audiences will struggle with this caveat to interpeting the map.  To demonstrate the possible consequences of this we treated the lower and upper quantile plots as though they were intensities and integrated them to obtain an expected abundance estimate of approximately 3,300 for the 0.025 quantile map and 12,400 for the 0.975 quantile map.  A naive (and tempting) interpretation of these numbers is as lower and upper limits of the 95\% credible interval for abundance. However, these abundance estimates are outside the support of the posterior for abundance (see \autoref{fig:realized-abundance-posterior}).  This demonstrates that the natural tendency to interpet these maps as possible intensities leads to interpretations that are inconsistent with the model that generated the maps.

The standard deviation map is a useful one but care has to be taken to think about what the standard deviation values might imply in the context of a given analysis.  A default colour scale will more or less always show some spatial variation in uncertainty with some regions of relatively high or low uncertainty.  What such a map does not show is whether these values actually matter.  Perhaps the posterior predictions across the entire region are all so uncertain as to the make the predictions operationally useless everywhere.  Or perhaps differences between high and low uncertainty regions (`high' and `'low' as defined by the colour scale) are small enough to be negligible when it comes to decision making, in which case such a spatially varying map is not a relevant output of the analysis.

Our solution to the problems in the quantile and standard deviation maps is to suggest analysts and stakeholders agree \emph{a priori} on values of the process and uncertainty thresholds that are important in the context of the analysis.  These values can then be used to construct summary maps.  We demonstrate this perspective through the use of excursions sets and functions \citep{bolin_excursion_2015}, although there are many possible approaches.  These techniques require the user to specify thresholds of interest for the process itself and acceptable levels of uncertainty.

This approach to representing variability in stochastic processes specifically considers the joint probability of events across multiple locations.  The positive excursion set with level $u$ for a function $f(s)$ with domain $\Omega$ is $A_u^{+}(f) = \{ s \in \Omega ; f(s) > u \}$ i.e. the set of all locations in $\Omega$ where $f$ exceeds the a threshold value $u$. For a stochastic process $\lambda(s)$ the positive excursion set with level $u$ and probability $1 - \alpha$ is
\begin{equation*}
E_{u,\alpha}^{+}(\lambda) = \argmax_{D}\{\lvert D \rvert : \mathbb{P}\left[D \subset A_u^{+}(\lambda)\right] \geq 1 - \alpha \}
\end{equation*}
Note that $A_u^{+}(f)$ specifies a set for which a function $f(s)$ exceeds a threshold value $u$ for \textit{every location} in the set.  As such the positive excursion set $E_{u,\alpha}^{+}(\lambda)$ is the largest such set for which a threshold is exceeded simultaneously for all locations in the set.  Negative excursion sets can be similarly defined.  Excursion sets can be estimated by considering candidate sets for $D$ of increasing size and a sequential integration scheme to estimate probabilities.  An implementation is available in the \texttt{excursions} package \citep{bolin_calculating_2018} available through the Comprehensive R Archive Network \citep{r_2017}.

\autoref{fig:excursions} (left panel) shows the positive excursion set with a level corresponding to 1 bird per hectare with probability 0.95.  This figure can be interpreted in a natural way.  It is the largest region for which the intensity is greater than 1 bird per hectare for every location within the region, with probability 0.95.
\begin{figure}
	\includegraphics[scale=0.5]{figures/excursions.png}
	\caption{Left:  The positive excursion set with a level corresponding to 1 bird per hectare and probability 0.95.  Right: The positive excursion function with a level corresponding to 1 bird per hectare}
	\label{fig:excursions}
\end{figure}
To visualise multiple such maps we can use the excursion function $F_u^{+}(s) = \sup \{1 - \alpha ; s \in E_{u,\alpha}^+ \}$, which defines for each location something similar to a p-value.  For each $s$ it is the largest possible probability $1 -\alpha$ for which that location would be in the excursion set defined using probability $1 - \alpha$.  The excursion function with a level corresponding to 1 bird per hectare is shown in \autoref{fig:excursions} (right panel).  This figure can also be interpreted naturally.  It shows the largest possible probability for each location to be a member of a set in which the intensity exceeds a threshold simultaneously across all locations in the set. It is clear from the figure that regions on the edge of the excursion set would be included if the $\alpha$ value were allowed to increase slightly.  However, for regions in the north there is essentially no probability level for which those locations would be included in $A_u^{+}(\lambda)$.  
We chose the threshold of 1 bird per hectare and an error probability of 0.05 to demonstrate the approach.  Multiple values for each of these can be considered as a simple extension.

Our intention in this paper was to present a new approach to analysing point transect distance sampling data that incorporates several innovative new methods and considers the entire workflow of model specification, inference, evaluation and communication of results.  Although each of these areas are substantive enough to be the main focus of a paper, we hope our brief presentation of issues that arise in each area and our chosen solutions will be useful to statisticians who face similar questions and data structures.

spatio-temporal extension. AR1 through time etc.  Citations for this.

This difference between summaries of the posterior and possible intensity surfaces that are consistent with the data becomes clear if one generates realizations of the posterior intensity field.  Three such realizations are in \autoref{fig:intensity-realizations}.  There are some notable differences when compared to the mean and quantile summary maps.  Each realization has finer-grained spatial structure than the mean or quantile plots of the posterior intensity.  In practice, this indicates that we should expect more spatial structure in our collected data than what we might expect if we only look at the posterior mean of the field.  The summary plots hide the spatial structure the model has been able to estimate from the data.  In general, the mean of a random field will contain less spatial structure than realisations of the field.

@@
